How to improve your relationship with your future self.

### 2011-03-22 workflow.Rnw
More writing

Using R, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called
\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
  R functions used in this essay come from the
  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
  data analysis. The short hand commands like \Verb+please-plot+ can
  be blocked and \Verb+may-I-please-have-a-plot+ can be required using
  \Verb+options(politeness=99)+ }

\begin{Verbatim}[fontsize=\small]
thedata<-read.csv("Data/thedata.csv")
pdf('fig1.pdf')
please-plot(outcome by explanatory using thedata. red lines please.)
dev.off()
\end{Verbatim}

@@

\section{No data analyst is an island.}
Data analysis involves a long series of decisions. Each decision
requires justification.  The body of the paper or book or report will
justify the most important decisions. Some decisions will be too small
and technical for inclusion in the paper itself. These need to be
documented in the code itself \citep{nagler1995coding}.

Comments --- unexecuted text inside of a script --- are a message to
your future self and other consumers of your work.  Read
. Comment your code. Luckily, if you are using
a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code as well a
technical comments in the code itself. If you are not using a strictly
literate programming practice then use whatever commenting protocol
exists in your analysis language. Here are some examples:
\begin{verbatim}
# This is a comment in R.
# With a a second line.

* This is a comment in a Stata .do file
* You can also embed comments at the ends of lines using /* ... */

/* SAS uses a similar convention to Stata
   and allows multi-line comments like this.
*/

% Finally, the percent symbol is a comment in TeX.
\end{verbatim}

Using R, for example, the comment is marked using the
pound sign. Notice that I've left in a regression not run for the
paper directly, but run as a part of the due diligence process of
writing the paper.

<<reg1,echo=TRUE,results=verb>>=

 ##  Repeat the regression run for Table~\ref{tab:protest}
 ##  Do political rights predict protest reporting?
 lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
 ##  Q: Was it worthwhile controlling for political rights?
 lm2<-lm(protac2000~I(Gini2004/100),data=good.df)
 coef(lm2)[2]-coef(lm1)[2]
 ##  A: Yes (around 3/4 an act of protest difference)
@

If you need to track down something more than a small detail, but yet
the sleuthing is really a footnote to the article, then you can create
another file to explore and report on those questions for the research
group. The writing in such a memo can be more informal, but the norms
of reproducibility ought to be the same as for the main paper. After
all, you will be releasing both documents (probably) publicly upon
publication of the article if not before.

Notice one another benefit of coding for an audience: we learn by
teaching. By assuming that others will look at your code, you will be
more likely to write clearer code, or perhaps even to learn more about
what you are doing as you do it.

Comment liberally. Comments are discarded when R runs analysis or LaTeX turns
your poignant writings into a PDF, so only those who dig into your work will
see them. They are footnotes on the process of producing your research.

>>>>

Using R, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called
\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
  R functions used in this essay come from the
  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
  data analysis. Functions like \Verb+please-plot+ can
  be blocked and more polite versions such as \Verb+may-I-please-have-a-plot+ can be required using
  \Verb+options(politeness=99)+ }

\begin{Verbatim}[fontsize=\small]
thedata<-read.csv("Data/thedata-15-03-2011.csv") ## Read the data
pdf('fig1.pdf') ## begin writing to the pdf file
please-plot(outcome by explanatory using thedata. red lines please.)
dev.off() ## stop writing to the pdf file
\end{Verbatim}

@@

\section{No data analyst is an island for long.}
Data analysis involves a long series of decisions. Each decision
requires justification.  The text itself will
justify the most important decisions. Some decisions will be too small
and technical for inclusion in the paper itself. These need to be
documented in the code itself \citep{nagler1995coding}.

Comments --- unexecuted text inside of a script --- are a message to
your future self and other consumers of your work. Luckily, if you are
using a literate programming practice (ex. Sweave or odfWeave), you
can write paragraphs to surround your code as well as technical
comments in the code itself. If you are not using a strictly literate
programming practice then use whatever commenting protocol exists in
your analysis language. R, for example, considers text marked with
``\#'' as a comment. Here for example I comment about a linear model
run for a paper and also include comments about another model run for
due diligence.

<<reg1,echo=TRUE,results=verb>>=
##  Repeat the regression run for Table~\ref{tab:protest}
##  Do political rights predict protest reporting?
lm1 <- lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
##  Q: Was it worthwhile controlling for political rights?
lm2 <- lm(protac2000~I(Gini2004/100),data=good.df)
coef(lm2)[2]-coef(lm1)[2]
##  A: Yes (around 3/4 an act of protest difference)
@

Why leave messages for yourself (or others)? These messages help you
retrace and justify the steps of your data-analytic journey in case
you ever need to revist them as the work moves from seminar paper to
conference paper to poster back to paper to dissertation and
onwards. Notice one another benefit of coding for an audience: we
learn by teaching. By assuming that others will look at your code, you
will be more likely to write clearer code, or perhaps even to learn
more about what you are doing as you do it.

Sometimes simple comments in the main command file are not enough:
imagine if the authors had arrived at the bootstrap only after running
simulations to assess the coverage of that procedure compared to a
profiled likelihood based approach. Rather than include time consuming
simulations in the main body of their Sweave document, they might have
written some memos that discuss the tradeoffs and implementation
decisions involved in such a comparison. In the main text they would
have just read in the results of the simulation study with an
appropriate comment:

<<bsreg1,echo=TRUE,results=verb,eval=FALSE>>=
## We assessed the bootstrap against direct inspection of the likelihood profile
## in the file Memos/shouldwebootstrap.Rnw. Here we read in the results of the bootstrap procedure:
load("Memos/bsresults.rda") ## using a compressed binary format here for speed
@

Comment liberally. Comments are discarded when R runs analysis or
LaTeX turns text into a pdf, so only those who dig into your work will
see them. They are footnotes on the process of producing your
research.

% If you need to track down something more than a small detail, but yet
% the sleuthing is really a footnote to the article, then you can create
% another file to explore and report on those questions for the research
% group. The writing in such a memo can be more informal, but the norms
% of reproducibility ought to be the same as for the main paper. After
% all, you will be releasing both documents (probably) publicly upon
% publication of the article if not before.
